Gyrochronal and Lithium Ages for Stars and Planets Observed by Kepler

### 2023-06-06 paper/ms.tex
update ms

\begin{abstract}
  We report rotation- and lithium-based ages for stars and planets observed by
  the primary Kepler mission.  Stellar rotation periods are collected from
  previous analyses of the Kepler light curves; lithium is assessed directly from HIRES
  spectra.
  wWe report ages for \nkoisgoldwithages\ FGK stars with transiting planets, with relative statistical
  precisions $\sigma_t / t$ of XX +YY -ZZ\%; a factor of X.X
  +Y.Y -Z.Z more precise than state-of-the-art isochronal ages for the
  same stars.  The results include XXX planets younger than 1\,Gyr,
  and YYY planets younger than 0.5\,Gyr.
  In the base sample of dwarf FGK stars observed by
  Kepler, we find that the intrinsic age distribution is(/not)
  consistent with a uniform distribution between X and Y billion years
  (Gyr).  In the planet sample, we find a consistent(/inconsistent)
  distribution at $N$-sigma.  Analyzing the completeness of the Kepler
  detection pipeline, we interpret the majority of the change to be
  caused by selection effects (or astrophysics).  We also comment on a
  few previously reported evolutionary trends in the size evolution of
  close-in mini-Neptunes (2-4\,\rearth) and super-Earths
  (1-2\,\rearth), a few of which we recover at comparable
  ($\sim$3$\sigma$) significance to previous work, and one of which
  we note for the first time.
\end{abstract}

@@

Concatenating the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} yields \nstarswithprot\ Kepler stars with
rotation periods.
%FIXME

@@

We considered two different samples of planets: those in the fully
automated Q1-Q17 DR25 KOI Table \citep{Thompson_2018}, which are
suitable for planet occurrence rate calculations, and also those in
the cumulative KOI table, which includes the best knowledge available
on any given planet candidate, but incorporates human-based vetting
\footnote{See
\url{https://exoplanetarchive.ipac.caltech.edu/docs/PurposeOfKOITable.html}}.
For the main body of this work, we will refer to the former
homogeneous sample, which includes \nkois\ KOIs, \nkoisnofp\ that are
not known false positives;
Appendix~\ref{app:cumulativekoi} highlights some additional young
planets that are omitted in this approach.

@@

>>>>

\begin{abstract}
  We report rotation- and lithium-based ages for stars and planets observed by
  the primary Kepler mission.  Stellar rotation periods are collected from
  previous analyses of the Kepler light curves; lithium content is assessed
  using HIRES spectra.
  We report ages for \nkoisgoldwithages\ FGK stars with transiting planets, with relative statistical
  precisions $\sigma_t / t$ of XX +YY -ZZ\%; a factor of X.X
  +Y.Y -Z.Z more precise than state-of-the-art isochronal ages for the
  same stars.  The results include XXX planets younger than 1\,Gyr,
  and YYY planets younger than 0.5\,Gyr.
  In the base sample of dwarf FGK stars observed by
  Kepler, we find that the intrinsic age distribution is(/not)
  consistent with a uniform distribution between X and Y billion years
  (Gyr).  In the planet sample, we find a consistent(/inconsistent)
  distribution at $N$-sigma.  Analyzing the completeness of the Kepler
  detection pipeline, we interpret the majority of the change to be
  caused by selection effects (or astrophysics).  We also comment on a
  few previously reported evolutionary trends in the size evolution of
  close-in mini-Neptunes (2-4\,\rearth) and super-Earths
  (1-2\,\rearth), a few of which we recover at comparable
  ($\sim$3$\sigma$) significance to previous work, and one of which
  we note for the first time.
\end{abstract}

@@

Concatenating the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} yields \nstarswithprot\ Kepler stars with
rotation periods.

%FIXME

@@

We considered two different samples of planets: those in the fully
automated Q1-Q17 DR25 KOI Table \citep{Thompson_2018}, which is
suitable for planet occurrence rate calculations, and also those in
the cumulative KOI table, which included the best knowledge available
on any given planet candidate as of 2023 Jun 6, but incorporated
human-based vetting \footnote{See
\url{https://exoplanetarchive.ipac.caltech.edu/docs/PurposeOfKOITable.html}}.
For the main body of this work, we will refer to the former
homogeneous sample, which includes \nkois\ KOIs, \nkoisnofp\ that are
not known false positives;
Appendix~\ref{app:cumulativekoi} highlights some additional young
planets that are omitted in this approach.

@@

As a sanity check on the statistical uncertainties of these rotation
periods, we compared our adopted \citetalias{Santos_2019} and
\citetalias{Santos_2021} periods with those reported by
\citet{McQuillan_2014} and \citet{Mazeh_2015}.
We found that for $P_{\rm rot}\lesssim$15\,days, the two 
datasets at a precision of $\lesssim$0.01$P_{\rm rot}$.  At longer periods of 
$P_{\rm rot}\approx$30\,days, the agreement was typically at the
$\lesssim$0.03$P_{\rm rot}$ level, and
the upper envelope of the period difference tended to increase
linearly with period.
Based on this comparison, we adopted a simple prescription for the
period uncertainties,
such that there are 1\% relative uncertainties below $P_{\rm
rot}=15$\,days, and a linear increase thereafter, with slope set to
guarantee 3\% $P_{\rm rot}$ uncertainties at 30 days.
